---
title: "Hwk5_bs2996"
author: "Bangda Sun"
date: "October 25, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Data for the US

### i. Write function: percentile_ratio_discrepancies.

```{r}
percentile_ratio_discrepancies <- function(P99, P99.5, P99.9, a){
  # input P99, P99.5, P99.9
  # return the sum square
  sse <- ((P99 / P99.9)^(-a+1) - 10)^2 + 
         ((P99.5 / P99.9)^(-a+1) - 5)^2 + 
         ((P99 / P99.5)^(-a+1) - 2)^2
  return(sse)
}
# Check the return
percentile_ratio_discrepancies(P99 = 1e6, P99.5 = 2e6, P99.9 = 1e7, a = 2)
```

### ii: Write function: exponent.multi_ratios_est

```{r}
exponent.multi_ratios_est <- function(P99, P99.5, P99.9){
  # input P99, P99.5, P99.9
  # minimize the percentile_ratio_discrepancies using nlm()
  # initial value a determined by formula (4)
  a   <- (1 - (log(10) / log(P99 / P99.9)))
  est <- nlm(percentile_ratio_discrepancies, a, P99 = P99, P99.5 = P99.5, P99.9 = P99.9)
  return(est$estimate)
}
exponent.multi_ratios_est(P99 = 1e6, P99.5 = 2e6, P99.9 = 1e7)
```


### iii: Write function and plotting the exponential estimates.

```{r}
# set work directory and import data
setwd("C://Users//Bangda//Desktop//GR5206 Materials//Hwk5")
wtid <- read.csv("wtid-report.csv", header = TRUE)
str(wtid)
names(wtid) <- c("Country", "Year", "P99", "P99.5", "P99.9")
# function a.estimate 
a.estimate <- function(P99, P99.5, P99.9){
  # input P99, P99.5, P99.9
  # using exponent.multi_ratios_est to estimate each years' a
  a1 <- rep(NA, length(P99))
  for (i in 1:length(a1)){
    # if one of P99, P99.5, P99.9 is missing, return NA
    if (is.na(P99[i]) | is.na(P99.5[i]) | is.na(P99.9[i])){
      a1[i] <- NA
    } else {
      # if P99, P99.5, P99.9 all exists, estimate a
      a1[i] <- exponent.multi_ratios_est(P99 = P99[i], P99.5 = P99.5[i], P99.9 = P99.9[i])
    }
  }
  return(a1)
}
a1 <- a.estimate(P99 = wtid$P99, P99.5 = wtid$P99.5, P99.9 = wtid$P99.9)
df1 <- data.frame(Year = wtid$Year, a1 = a1)
library(ggplot2)
ggplot(df1, mapping = aes(x = Year, y = a1)) + geom_line() + 
  labs(x = "Year", y = "Estimation of a")
```

### iv. Comparison of two estimation

```{r}
# using equation (4) to estimate
a2 <- (1 - (log(10)) / (log(wtid$P99 / wtid$P99.9))) 
df2 <- data.frame(a1 = a1, a2 = a2)
ggplot(df2, mapping = aes(x = a1, y = a2)) + geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  labs(x = "Estimation of a using a.estimate", y = "Estimation of a using equation (4)")
identical(a1, a2)
```

These two estimates are highly correlated but not totally identical.


## Part 2: Data for Other Countries

### v. Estimate a for each country over time.
```{r}
library(xlsx)
# import the .xlsx file and save them into data.list
# data.list should has 9 data frame to store 9 countries' data
data.files <- list.files(pattern = "*.xlsx")
data.list <- vector("list", length(data.files))
for (i in 1:length(data.files)){
  data.list[[i]] <- read.xlsx(data.files[i], sheetName = "Series-layout A", 
                              startRow = 2, header = TRUE) 
}

# convert the all countries' data in a single dataframe
wtid3 <- data.frame(Country = character(0), Year = numeric(0),
                    AveIncomePerTU = numeric(0), P99 = numeric(0),
                    P99.5 = numeric(0), P99.9 = numeric(0))
for (j in 1:length(data.files)){
  # check if the variables exist in the dataframe
  if("P99.income.threshold" %in% colnames(data.list[[j]]) & 
     "P99.5.income.threshold" %in% colnames(data.list[[j]]) & 
     "P99.9.income.threshold" %in% colnames(data.list[[j]]) &
     "Average.income.per.tax.unit" %in% colnames(data.list[[j]])){
    # insert the value into the new dataframe
    wtid3 <- rbind(wtid3, data.frame(
      Country = data.list[[j]]$Country,
      Year = data.list[[j]]$Year,
      AveIncomePerTU = data.list[[j]]$Average.income.per.tax.unit,
      P99 = data.list[[j]]$P99.income.threshold,
      P99.5 = data.list[[j]]$P99.5.income.threshold,
      P99.9 = data.list[[j]]$P99.9.income.threshold
  ))
  } else {
    # if the variables are not in the dataframe, create them
    data.list[[j]]$Average.income.per.tax.unit <- rep(NA, length(data.list[[j]]$Year))
    data.list[[j]]$P99.income.threshold <- rep(NA, length(data.list[[j]]$Year))
    data.list[[j]]$P99.5.income.threshold <- rep(NA, length(data.list[[j]]$Year))
    data.list[[j]]$P99.9.income.threshold <- rep(NA, length(data.list[[j]]$Year))
    wtid3 <- rbind(wtid3, data.frame(
      # insert the value NA into the new dataframe
      Country = data.list[[j]]$Country,
      Year = data.list[[j]]$Year,
      AveIncomePerTU = data.list[[j]]$Average.income.per.tax.unit,
      P99 = data.list[[j]]$P99.income.threshold,
      P99.5 = data.list[[j]]$P99.5.income.threshold,
      P99.9 = data.list[[j]]$P99.9.income.threshold
    ))
  }
}
# estimate a for each country and each year
wtid3$a <- a.estimate(P99 = wtid3$P99, P99.5 = wtid3$P99.5, P99.9 = wtid3$P99.9)
```

### vi. Plotting estimate of a over time

```{r}
ggplot(data = wtid3) + 
  geom_line(mapping = aes(x = Year, y = a, col = Country))
```

### vii. Plotting average of income per tax unit

We can see that the average income per tax unit varies a lot among different countries, therefore we take logarithm to visualize them more clearly.

```{r}
library(gridExtra)
g1 <- ggplot(data = wtid3) + 
  geom_line(mapping = aes(x = Year, y = AveIncomePerTU, col = Country)) +
  labs(x = "Year", y = "Average income \nper tax unit")
g2 <- ggplot(data = wtid3) + 
  geom_line(mapping = aes(x = Year, y = AveIncomePerTU, col = Country)) +
  scale_y_log10() + 
  labs(x = "Year", y = "logarithm of Average income \nper tax unit")
grid.arrange(g1, g2, ncol=1)
```


### viii. Plotting a against average income for the US
```{r}
# subset US data
wtid.us <- wtid3[wtid3$Country == "United States", ]
# ploting
ggplot(data = wtid.us) +
  geom_point(mapping = aes(x = AveIncomePerTU, y = a))
```

The graph shows that $a$ is small at the very beginning, then it rises as average income rises (inequality decline), and then it goes down (inequality growth). Therefore Kuznets curve is not reasonable for the situation of US.

### ix. Building quadratic regression model for US
```{r}
lm1 <- lm(a ~ AveIncomePerTU + I(AveIncomePerTU^2), data = wtid.us)
summary(lm1)
```

As we can see, the coefficient of quadratic term is less than zero, which means the fitted function is a concave function, therefore the coefficient is not consistent with the hypothesis. 

### x. Building quadratic regression
```{r}
reg <- function(country){
  # check if any necessary data exists
  if (all(is.na(wtid3$a[wtid3$Country == country]))){
    return(cat("Data of", country,"is missing."))
  } else {
    lm <- lm(a ~ AveIncomePerTU + I(AveIncomePerTU^2), data = wtid3[wtid3$Country == country,])
    return(summary(lm))
  }
}
reg("Canada")
reg("China")
reg("Colombia")
reg("Germany")
reg("Italy")
reg("Japan")
reg("South Africa")
reg("Sweden")
```

As we can see, the coefficients of quadratic term is positive for these countries: China, Colombia and Japan. And the coefficients are only reliable for Japan. Therefore Japan is compatible with the hypothesis.

